---
title: "Exploring The Beautiful Game"
output: github_document
---


I came across [this soccer dataset](https://www.kaggle.com/hugomathien/soccer/data) and decided to explore it using R. 

We begin loading the data.
```{r}
library(DBI)
library(RSQLite)
library(dplyr)

# Database connection.
con <- dbConnect(SQLite(),"database.sqlite")

# Names of tables.
names <- as.data.frame(dbListTables(con))

# Tables.
player <- dbReadTable(con,'Player')
playerAttr <- dbReadTable(con,'Player_Attributes')

# Disconnecting from the database.
dbDisconnect(con)
```

Let's arrange the players by rating and find out who were the 50 players with the highest FIFA ratings during this period.
```{r}
# Returns the name of player p.
getName <- function(p) {filter(player,p$player_api_id==player_api_id)$player_name}

# Top 50 distinct player attributes ordered by FIFA ratings.
top50 <- distinct(na.omit(arrange(playerAttr,desc(overall_rating))),player_api_id,.keep_all=TRUE)[1:50,]

# Show the names of the 50 players with the highest rankings in descending order.
sapply(c(1:nrow(top50)), function(i) paste(i,"-",getName(top50[i,]),"\n")) %>% message
```

As expected, Messi and Ronaldo occupy positions 1 and 2.

Which of these players are over 190 centimeters tall?
```{r}
# Show players over 190 cm. tall.
filter(player, player_api_id%in%(top50$player_api_id) & height>190)[,-c(1,4,5)]
```

Which of these players weigh under 160 pounds?
```{r}
# Show players under 160 pounds.
filter(player, player_api_id%in%(top50$player_api_id) & weight<160)[,-c(1,4,5)]
```

Surprisingly, Robin van Persie is among these players.

What is the average height and weight of the top 50 highest ranked players?
```{r}
# Show mean height and weight.
summarise(filter(player,player_api_id%in%(top50$player_api_id)),mean(height),mean(weight))
```

What is the preferred foot ratio?
```{r}
# Show number of left and right footed players.
count(top50,preferred_foot)
```

Everything looks reasonable so far, but to further evaluate the quality of the data lets test the idea that similar players play at similar positions by grouping these players into four clusters (representing the four main positions in the game: goalkeeper, defender, midfielder, and striker).
```{r}
# K-means object.
clusters <- kmeans(top50[,10:42],centers=4,iter.max=20,nstart=100)

# Returns the names of the players of the nth cluster.
clusterNames <- function(n) {
   clust <- as.data.frame(split(top50,clusters$cluster)[n])
   names(clust) <- names(top50)
   sapply(c(1:nrow(clust)), function(i) getName(clust[i,]))
}

# Matrix of names of players per cluster.
playersPerCluster <- sapply(c(1:length(clusters$size)),clusterNames)

# Show names of players per cluster.
sapply(c(1:length(clusters$size)), function(i) (playersPerCluster[i]))
```

We can see that two of these clusters are composed exclusively by goalkeepers and defenders, while the other two also seem to distinguish between midfielders and strikers.