---
title: "Exploring The Beautiful Game"
output: html_notebook
---


I came across [this soccer dataset](https://www.kaggle.com/hugomathien/soccer/data) and decided to explore it using R. 

We begin loading the data.
```{r}
library(DBI)
library(RSQLite)
library(dplyr)

# Database connection
con <- dbConnect(SQLite(),"database.sqlite")

# Names of tables
names <- as.data.frame(dbListTables(con))

# Tables
player <- dbReadTable(con,'Player')
playerAttr <- dbReadTable(con,'Player_Attributes')

# Disconnecting from the database
dbDisconnect(con)
```

Let's order the players by rating and find out who were the 50 players with the highest FIFA ratings during this period.
```{r}
# Distinct player attributes ordered by FIFA ratings.
topRatings <- distinct(na.omit(arrange(playerAttr,desc(overall_rating))),player_api_id,.keep_all=TRUE)

# Returns the name of player p.
getName <- function(p) {filter(player,p$player_api_id==player_api_id)$player_name}

# Print the names of the 50 players with the highest rankings in descending order.
for (i in 1:50) {message(paste(i,"-",getName(topRatings[i,])))}
```

As expected, Messi and Ronaldo occupy positions 1 and 2.

Which of these players are over 190 centimeters tall?
```{r}
filter(player, player_api_id%in%(topRatings[1:50,]$player_api_id) & height > 190)
```

Which of these players weigh under 160 pounds?
```{r}
filter(player, player_api_id%in%(topRatings[1:50,]$player_api_id) & weight < 160)
```

Surprisingly, Robin van Persie is among these players.

What is the average height and weight of the top 50 highest ranked players?
```{r}
summarise(filter(player,player_api_id%in%(topRatings[1:50,]$player_api_id)),mean(height),mean(weight))

```

What is the preferred foot ratio?
```{r}
count(topRatings[1:50,],preferred_foot)
```

Everything seems reasonable so far, but to further test the data lets test the idea that similar players play at similar positions by grouping these players into four clusters representing the four main positions in the game (goalkeeper, defender, midfielder, and striker).
```{r}
# Lets create clusters of the players...
clusters <- kmeans(topRatings[1:50,10:42],centers=4,iter.max=20,nstart=100)

# Printing the names of players of each cluster.
for (n in 1:4) {
   c <- as.data.frame(split(topRatings[1:50,],clusters$cluster)[n])
   names(c) <- names(topRatings)
   message(paste("CLUSTER ",n,":"))
   for (i in 1:nrow(c)) {message(getName(c[i,]))}
   message("-----------------")
}
```

